---
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
---
title: "DSCC/CSC/STAT 462 Final Project"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: "Unmasking TV Series' Statistics"
fontsize: 12pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval=T,tidy=TRUE, tidy.opts=list(width.cutoff=60))
```
  
```{r}
source("./utils.R")
library(readxl)
```

```{r}
df <- read_excel("TMDB_tv_dataset_v3.xlsx")
print(colnames(df))
```

```{r}
# dropping non-useful columns
df <- subset(df, select = -c(id,overview,homepage,poster_path))
print(colnames(df))
dim(df)
```

Exploring numeric variables and their usability
```{r}
#NUMBER OF SEASONS
col_stats(df,"number_of_seasons")
```
 
by the summary statistics it is clear that for some of the TV shows,
the number of seasons has values as 0 (min), which implies that those shows
either did not follow the concept of seasons or the information is not available

Also since Q1=Q2=Q3 we can conclude that almost 75% of the values equal to 1 implying
one season only, however the maximum value is 240 which is significantly higher

The skewness value of 14.022 indicates substantial right skewness, i.e., extreme values in 
the higher range. 

The IQR has been reported as 0, suggesting that the middle 50% of the data points 
are relatively tightly clustered around the median (i.e., the middle 50% of values = 1)

```{r}
generate_box_plot(
  data = df, 
  y_col = "number_of_seasons", 
  y_label = "Number of seasons", 
  title = "Box plot of number of seasons of all TV shows",
  ylim = c(0, 20)
)
```

The box plot obtained shows that the middle 50% of the values are equal to 1 as the Q1, Q2 and Q3 
have converged to 1 whereas higher number of seasons have been shown as outliers in the plot.

```{r}
plot_histogram(
  data = df, 
  column_name = "number_of_seasons", 
  x_label = "Range of Season Count",
  y_label = "Number of TV Series",
  title = "Histogram of TV Series Season Count",
  type = 0,
  fillColor = "green")

```

The histogram further confirms the right skewness of the variable and that more than 100000 observations have 0 or 1 season. But since Q1 = Q2 = Q3 = 1, the majority of observations seem to have a single season only. 

```{r}
#NUMBER OF EPISODES
col_stats(df,"number_of_episodes")
generate_box_plot(
  data = df, 
  y_col = "number_of_episodes", 
  y_label = "Number of episodes", 
  title = "Box plot of the number of episodes of TV series",
  ylim = c(0, 100)
)
```

Inference: 
The minimum and maximum values being 0 and 20839 indicate a wide range, suggesting 
significant variability in the number of episodes. The mean of 25.06 suggests average
number of episodes but the presence of high outliers seem to influence the value.
Median of 6 is considerably lower than the mean, suggesting positively skewed distribution
due to the influence of high values. Skewness of 54.38 indicates significant positive 
skewness. The 10% trimmed mean of 10.51, calculated by removing extreme values, is lower 
than the mean and likely reflects a more representative average in the presence of 
outliers. 

```{r}
#HISTOGRAM  
plot_histogram(
  data = df, 
  column_name = "number_of_episodes", 
  x_label = "Range of Episode Count",
  y_label = "Number of TV Series",
  title = "Histogram of TV Series Episode Count",
  type = 0,
  fillColor = "skyblue")
  
#QQ PLOT
generate_qq_plots(data = df, column_name = "number_of_episodes")
```

As further evident from the histogram and qq-plot, the number_of_episodes column is highly right
skewed and non-normal and therefore, we can apply the box-cox transformation to make it 
closer to normal.

```{r}
#BOX-COX TRANSFORMATION
transformed_col_df <- apply_boxcox_transformation(df, "number_of_episodes")
cat(str(transformed_col_df$transformed_number_of_episodes))
colnames((transformed_col_df))

generate_qq_plots(data = transformed_col_df, column_name = "transformed_number_of_episodes")

```

As we can see from quantile plot, the data for number_of_episodes is now much closer
to a normal distribution.
Further preprocessing of the variable (example taking >0 or >1 or exluding outliers)
can be done depending upon the test required. 

```{r}
df <- transformed_col_df

#VOTE AVERAGE
col_stats(df,"vote_average")
generate_box_plot(
  data = df, 
  y_col = "vote_average", 
  y_label = "Vote average (Rating)", 
  title = "Box plot of the vote average (ratings) of TV shows",
  ylim = c(0, 100)
)

```

The minimum value of 0 and maximum of 10 indicate that vote_average represents ratings on a scale of 0 to 10. The mean of 2.409 is notably lower than the median of 0.000, suggesting a right-skewed distribution with potential outliers pulling the mean upwards.
The median of 0 suggests that a considerable portion of the dataset consists of values at or close to zero. A skewness value of 0.914 indicates a moderately positive skewness, suggesting a tail on the right side of the distribution. The IQR of 6 demonstrates the range of the middle 50% of the data, which spans from 0 to 6. The 10% trimmed mean of 1.888, calculated by removing extreme values, is lower than the mean and might offer a more representative average in the presence of outliers. 

```{r}
#HISTOGRAM
plot_histogram(
  data = df, 
  column_name = "vote_average", 
  x_label = "Vote Average",
  y_label = "Number of TV Series",
  title = "Histogram of TV Series Vote Average",
  type = 0,
  fillColor = "light green")
```

From the histogram and summary statistics it is evident that majority of the data has a vote_average value of 0 which seems like a null value and not part of the metric for rating. Therefore, assuming that vote_average (or the rating) ranges from 1 to 10. We will drop the rows with 0 as the vote_average.

```{r}
df <- subset(df, vote_average > 0)
dim(df)

#HISTOGRAM
plot_histogram(
  data = df, 
  column_name = "vote_average", 
  x_label = "Vote Average",
  y_label = "Number of TV Series",
  title = "Histogram of TV Series Vote Average",
  type = 0,
  fillColor = "light green")

#QQ PLOT
generate_qq_plots(data = df, column_name = "vote_average")
```

The histogram shows negative skewness and the quantile plot doesn't show the distribution to be normal, therefore, we can apply the boxcox transformation to normalize the distribution of vote_average

```{r}
#BOX-COX TRANSFORMATION
transformed_col_df <- apply_boxcox_transformation(df, "vote_average")
cat(str(transformed_col_df$transformed_vote_average))
colnames((transformed_col_df))

generate_qq_plots(data = transformed_col_df, column_name = "transformed_vote_average")
```

```{r}
df <- transformed_col_df
colnames(df)

#HISTOGRAM AFTER TRANSFORMATION
plot_histogram(
  data = df, 
  column_name = "transformed_vote_average", 
  x_label = "Vote Average",
  y_label = "Number of TV Series",
  title = "Histogram of TV Series Vote Average",
  type = 0,
  fillColor = "light green")
```

The histogram looks approximately normal, and therefore vote_average can be used a metric for analysing the TV shows data in combination with other variables. 


```{r}
#POPULARITY
col_stats(df,"popularity")
generate_box_plot(
  data = df, 
  y_col = "popularity", 
  y_label = "Popularity", 
  title = "Box plot of the popularity of TV shows",
  ylim = c(0, 1000)
)

```

The boxplot shows a heavily right skewed distribution with most of the outliers above the upper limit. The summary statistics indicate that a few items or instances have very high popularity scores (example the max value - 3707), which significantly impact the mean value. The majority of items might have lower popularity scores, while some outliers possess extremely high scores, widening the range (high variance) and skewing the distribution to the right. 

To be able to use popularity for analyses, we have to preprocess the attribute for outliers.

```{r}
data <- na.omit(df)

Q1 <- quantile(data$popularity, 0.25)
Q3 <- quantile(data$popularity, 0.75)
IQR <- Q3 - Q1

lower_limit <- Q1 - 1.5 * IQR
upper_limit <- Q3 + 1.5 * IQR

data <- subset(data, popularity > lower_limit & popularity < upper_limit)

data <- data.frame(data)

col_stats(data,"popularity")
generate_box_plot(
  data = data, 
  y_col = "popularity", 
  y_label = "Popularity", 
  title = "Box plot of the popularity of TV shows",
  ylim = c(0, 30)
)

df <- data.frame(data)
```

```{r}
# HISTOGRAM
plot_histogram(
  data = df, 
  column_name = "popularity", 
  x_label = "Popularity",
  y_label = "Number of TV Series",
  title = "Popularity of TV Series",
  type = 0,
  fillColor = "light blue")

# QQ PLOT
generate_qq_plots(data = df, column_name = "popularity")

```

After removing the outliers, the histogram and quantile plots indicate a right 
skewed distribution, however, applying the box-cox transformation here gives inappropriate values for popularity including negative and null values, therefore we do not apply transformation to popularity data. 

```{r}
#EPISODE RUN-TIME
col_stats(df,"episode_run_time")
generate_box_plot(
  data = df, 
  y_col = "episode_run_time", 
  y_label = "Total Runtime of episodes", 
  title = "Box plot of total episode runtimes of a scripted TV series",
  ylim = c(0, 500)
)
```

Due to the large variance, it is evident that some of the values in episode_run_time has been given in seconds and some in minutes. We therefore have to preprocess the attribute before analysis. Therefore, a 

This function iterates through the "episode_run_time" column of a data frame and performs the following steps:
1. Remove rows with missing values in "episode_run_time".
2. Update values outside the specified range directly.
3. Update values by dividing by "number_of_episodes" where applicable.
4. Update values by dividing by 60 where applicable.
5. Assign a default value of -1 for values that don't meet the conditions.
6. Remove rows with the default value.
7. Discard entire rows for values that don't meet the conditions.

We are going to take the limit for episode length in minutes to be a minimum of 10 minutes and a maximum of 90 minutes.

```{r}
df <- process_episode_data(df,10,90)
dim(df)

col_stats(df,"episode_run_time")
generate_box_plot(
  data = df, 
  y_col = "episode_run_time", 
  y_label = "Total Runtime of episodes", 
  title = "Box plot of total episode runtimes of a scripted TV series",
  ylim = c(0, 100)
)

```

Inference:

The mean of 42.24 minutes represents the average episode run time.
The median of 44 minutes suggests that half of the episodes have a duration of less than equal to 44 indicating a moderately symmetric distribution. Q1 = 27 and Q3 = 52 indicates that middle 50% of the episode run times lie between these values. With a coefficient of variation of 0.42, the standard deviation relative to the mean is relatively low, indicating moderate variability around the mean episode run time. A skewness value of 0.55 indicates a moderate positive skewness, suggesting a slightly longer tail on the right side of the distribution compared to a perfectly symmetric distribution. Overall, the summary statistics and the box plot suggest a distribution of episode run times that is moderately skewed to the right, but nearly normal with most episodes falling within the 27 to 52-minute range. While the mean and median are relatively close, the positive skewness indicates that some episodes might have durations longer than the median, contributing to the longer tail on the right side of the distribution. 

Exploring the categorical variables and their usability

Exploring genres in TV shows data
Since the genres column has multiple categories mentioned on same observations, we can expand the distribution based on comma separated values. This is done internally in the function call for plotting by calling the expand_nested_data function (defined in the utils.R file) 

```{r}
#GENRES BAR PLOT
plot_histogram(
  data = df, 
  column_name = "genres",
  x_label = "Genre",
  y_label = "Number of TV Series",
  title = "Genres of TV Series",
  type = 1,
  fillColor = "dark magenta")
```

The bar plot of different genres has been plotted above and it is evident from the plot that the most common genre among TV shows is "Drama", followed by "Comedy". In contrast, genres like "History", "Music", "Musical", etc. are the least common. 

```{r}
#ABSOLUTE FREQUENCY TABLE 
df_expanded_genres <- expand_nested_data(df,"genres")
colnames(df_expanded_genres)

colnames(df_expanded_genres)[colnames(df_expanded_genres) == "cat_col"] <- "genre"

for (i in 1:nrow(df_expanded_genres)) {
  print(paste(df_expanded_genres[i, ], sep = " "))
}
```

```{r}

#RELATIVE FREQUENCY TABLE
total_frequency <- sum(df_expanded_genres$frequency)

for (i in 1:nrow(df_expanded_genres)) {
  genre <- df_expanded_genres[i, "genre"]
  freq <- df_expanded_genres[i, "frequency"]
  percentage <- (freq / total_frequency) * 100
  
  # Print genre and relative frequency
  print(paste("Genre:", genre, "- RF :", round(percentage, 3), "%"))
}

```


Exploring "type" variable
```{r}
print(table(df$type))
```

Since the categories are not in comma separated, repeating format, there is no requirement of expansion, we can therefore get the frequency tables and bar plot directly

```{r}
#TYPE BAR PLOT
plot_histogram(
  data = df, 
  column_name = "type",
  x_label = "Type",
  y_label = "Number of TV Series",
  title = "Types of TV Series",
  type = 2,
  fillColor = "purple")
```

From the barplot of "type" categorical variable, we can observe that most of the TV shows listed in the data are of "Scripted" type. It is followed by "Miniseries", "Documentary" and "Reality". "News" and "Video" are the least common type categories. 

Frequency tables for categorical variable type:

```{r}
#ABSOLUTE FREQUENCY TABLE
freq_table_type <- sort(table(df$type), decreasing = TRUE)
print("Type  Frequency")
for (val in unique(df$type)) {
  print(paste(val, " ", freq_table_type[val]))
}
```

```{r}
#RELATIVE FREQUENCY TABLE
relative_freq_type <- prop.table(freq_table_type) * 100

print("Type  Relative-Frequency")
for (val in names(relative_freq_type)) {
  print(paste(val, " ", round(relative_freq_type[val], 2), "%"))
}
```

Exploring "original_language"
```{r}
print(unique(df$original_language))
```

The original languages of the TV shows are in abbreviated format, therefore we can map them to their full versions, as close as possible with information available on the internet:

```{r}
# Abbreviations list
abbreviations <- c("en", "es", "ko", "de", "ja", "sv", "th", "it", "zh", "da", "fr", "pt", "tr", "ru", "is","he", "no", "pl", "nl", "hi", "tl", "fi", "lb", "gl", "uk", "la", "ar", "cs", "bg", "hu","cy", "el", "ta", "sr", "cn", "fa", "bn", "ca", "ro", "id", "xx", "ur", "sk", "ms", "sh","af", "te", "hr", "vi", "si", "ga", "et", "ka", "sq", "az", "lv", "eu", "nb", "sl", "kn","mk", "bs", "zu", "lt", "jv", "so", "pa", "ne")

# Full names list corresponding to abbreviations
full_names <- c("English", "Spanish", "Korean", "German", "Japanese", "Swedish", "Thai", "Italian", "Chinese", "Danish","French", "Portuguese", "Turkish", "Russian", "Icelandic", "Hebrew", "Norwegian", "Polish", "Dutch", "Hindi","Tagalog", "Finnish", "Luxembourgish", "Galician", "Ukrainian", "Latin", "Arabic", "Czech", "Bulgarian", "Hungarian","Welsh", "Greek", "Tamil", "Serbian", "Chinese", "Persian", "Bengali", "Catalan", "Romanian", "Indonesian","Unknown", "Urdu", "Slovak", "Malay", "Serbo-Croatian", "Afrikaans", "Telugu", "Croatian", "Vietnamese", "Sinhalese","Irish", "Estonian", "Georgian", "Albanian", "Azerbaijani", "Latvian", "Basque", "Norwegian Bokmål", "Slovenian", "Kannada","Macedonian", "Bosnian", "Zulu", "Lithuanian", "Javanese", "Somali", "Punjabi", "Nepali")


language_map <- setNames(full_names, abbreviations)
print(language_map)
```

```{r}
#BAR PLOT
plot_histogram(
  data = df, 
  column_name = "original_language", 
  x_label = "Language",
  y_label = "Number of TV Series",
  title = "Original Language(ISO 639-1 Code) of TV Series",
  type = 2,
  fillColor = "darkmagenta")
```

From the bar plot of "original_language" of TV shows data, we can observe that majority of the TV shows available in the dataset are in "English", followed by "French" and "Spanish"

```{r}
#ABSOLUTE FREQUENCY TABLE
freq_table_ol <- sort(table(df$original_language), decreasing = TRUE)
print("Original_language  Frequency")
for (val in unique(df$original_language)) {
  print(paste(language_map[[val]], " ", freq_table_ol[val]))
}
```

```{r}
#RELATIVE FREQUENCY TABLE
relative_freq_ol <- prop.table(freq_table_ol) * 100

print("Type  Relative-Frequency")
for (val in names(relative_freq_ol)) {
  print(paste(language_map[[val]], " ", round(relative_freq_ol[val], 2), "%"))
}
```

Exploring "networks" variable for TV shows
```{r}
print(unique(df$networks)[1:20])
```

Since some of the unique values in "networks" column contains multiple networks separated by "," we can expand in the same way as done in genres. This is done internally in the function call for plotting by calling the expand_nested_data function (defined in the utils.R file)

```{r}
#NETWORKS BAR PLOT
plot_histogram(
  data = df, 
  column_name = "networks",
  x_label = "Networks",
  y_label = "Number of TV Series",
  title = "Different networks of TV Series",
  type = 1,
  fillColor = "darkgreen")
```

From the barplot of networks, we can observe that most common network in the TV shows dataset is "Netflix", followed by "Youtube","TV Globo" and "Prime Video"

```{r}
#ABSOLUTE FREQUENCY TABLE 
df_expanded_networks <- expand_nested_data(df,"networks")
colnames(df_expanded_networks)

colnames(df_expanded_networks)[colnames(df_expanded_networks) == "cat_col"] <- "network"

for (i in 1:nrow(df_expanded_networks)) {
  print(paste(df_expanded_networks[i, ], sep = " "))
}
```

```{r}

#RELATIVE FREQUENCY TABLE
total_frequency_networks <- sum(df_expanded_networks$frequency)

for (i in 1:nrow(df_expanded_networks)) {
  network <- df_expanded_networks[i, "network"]
  freq <- df_expanded_networks[i, "frequency"]
  percentage <- (freq / total_frequency_networks) * 100
  
  # Print genre and relative frequency
  print(paste("Network:", network, "- RF :", round(percentage, 3), "%"))
}

```

Exploring "status" variable
```{r}
print(table(df$status))
```

Since the categories are not in comma separated, repeating format, there is no requirement of expansion, we can therefore get the frequency tables and bar plot directly.

```{r}
#STATUS BAR PLOT
plot_histogram(
  data = df, 
  column_name = "status",
  x_label = "Status",
  y_label = "Number of TV Series",
  title = "Status of TV Series",
  type = 2,
  fillColor = "darkgreen")
```

Frequency tables for categorical variable 'status':

```{r}
#ABSOLUTE FREQUENCY TABLE
freq_table_status <- sort(table(df$status), decreasing = TRUE)
print("Status  Frequency")
for (val in unique(df$status)) {
  print(paste(val, " ", freq_table_status[val]))
}
```

```{r}
#RELATIVE FREQUENCY TABLE
relative_freq_status <- prop.table(freq_table_status) * 100

print("Status  Relative-Frequency")
for (val in names(relative_freq_status)) {
  print(paste(val, " ", round(relative_freq_status[val], 2), "%"))
}
```

Now, to further understand the relationships between variables, we can use visualisations to further 

```{r}
# Side-by-side boxplots for number of episodes by genre for TV series

expanded_df <- expand_categorical_cols(df,"genres","number_of_episodes")

generate_box_plot(
  data = expanded_df,
  fill_column = "cat_col",
  y_col = "num_col",
  y_label = "Number of episodes", 
  title = "Box plot of number of a episodes of TV series for various genres",
  ylim = c(0, 200)
)

```

Inference:
From the boxplots above, it can be said that the variability in the number of episodes is comparable between the different genres. The "Family" and "Kids" genres tend to have higher medians for number of episodes whereas "Documentary" tends to have the least. 

We can now check side-by-side box plots for vote_average (ratings) for different genres

```{r}
# Side-by-side boxplots for vote_average by genre for TV series

expanded_df <- expand_categorical_cols(df,"genres","vote_average")

generate_box_plot(
  data = expanded_df,
  fill_column = "cat_col",
  y_col = "num_col",
  y_label = "Vote average (ratings)",
  title = "Box plot of vote average of TV series for various genres",
  ylim = c(0, 25)
)

```

From the side-by-side boxplots it is evident that the variability of ratings (vote_average) in different genres is comparable. However there appears to be many outliers beyond the lower whisker. The plot is ideal for ANOVA test for comparison of means, which will be further explored in the inferential statistics section.

To compare the variability in popularity of "Japanese" vs "English" animated tv series, we can use side-by-side box plots

```{r}
# Side-by-side boxplots for popularity by original language for animated  TV series



generate_box_plot(
  data = subset(df, grepl("Animation", genres)),
  fill_column = "original_language",
  categories = c("en", "ja"),
  y_col = "popularity", 
  y_label = "Popularity", 
  title = "Box plot of popularity of animated TV series for English and Japanese",
  ylim = c(0, 50)
)

```

From the box-plots above, we can infer that the variablity in the distribution between animated TV shows in English is more than the variability of animated TV shows in Japanese. The median of animated shows in English is also higher (around 10) than the median of animated shows in Japanese (around 7).

We can now use scatter plots to understand the relationships between numerical variables. 

```{r}
# SCATTER PLOT BETWEEN POPULARITY AND VOTE_AVERAGE

# COVARIANCE 
covariance <- cov(df$popularity, df$vote_average)
print(covariance)

display_scatter_plot(
  data = df, 
  x_col = "popularity", 
  y_col = "vote_average",
  z_col = "type", 
  x_label = "Popularity", 
  y_label = "Average Vote", 
  title = "Scatter plot of average votes of a scripted TV series with popularity", 
  xlim = c(0, 45), 
  ylim = c(0, 10)
  )

```

The scatter plot of popularity vs vote_average depicts that for a wide range of "popularity" values the "vote_average" for many points remain the same, i.e., vote_average is constant for different values of Popularity in multiple cases. This indicates that the two variables are independent of each other. The covariance between the two variables is 0.848 (approximately). The color coding indicates that majority of the tv shows fall in the "scripted" type. 

```{r}
# SCATTER PLOT BETWEEN NUMBER OF SEASONS AND DURATION

# COVARIANCE 
new_df <- get_duration(df)
covariance <- cov(new_df$number_of_seasons, new_df$duration)
print(covariance)

display_scatter_plot(
  data = new_df,
  x_col = "number_of_seasons",
  y_col = "duration",
  z_col = "type",
  x_label = "Number of seasons",
  y_label = "Duration",
  title = "Scatter plot of number of seasons vs duration of TV Series", 
  #xlim = c(0, 45), 
  #ylim = c(0, 10)
  )

```

The scatter plot above shows that the number of seasons is constant for a wide range of values of duration of TV Series implying that the two variables are independent of each other. A visible pattern of the outliers showing linear relationship cannot be accounted for as majority points do not follow the pattern.




